On the entropy of isothermal dusty gas: application to protoplanetary disks

### 2016-08-05 summary.tex
minor changes
more discussions

\subsection{Dusty analogs of other gaseous instabilities}  
%gi
%rwi

>>>>

\subsection{Dusty analogs of other gaseous instabilities}

%gi
%rwi

\subsubsection{Gravitational instability} %vertical structure 
The addition of dust enhances gravitational
instability (GI). This is because dust particles contribute to the
total disk mass but not thermal pressure, which effectively lowers the
disk temperature \citep[][]{thompson88}. For typical dust-loading
$\epsilon, Z\ll1$, this effect is unimportant. However, if $\epsilon$ is
large (e.g. due to dust settling) then the effective temperature may 
be lowered to enable instability.

As noted in \S\ref{loc_iso_eos}, dust settling causes the effective
temperature to \emph{increase} away from the midplane. This contrasts
to previous studies of GI in vertically stratified disks
\citep[e.g.][]{mamat10, kim12,lin14c} where the temperature decreases
from the midplane. While we expect only the total surface density and
characteristic temperatures are relevant to stability 
\citep{toomre64}, having a non-trivial vertical temperature
structure, induced by dust, may modify the vertical structure of 3D
waves and unstable modes. We will examine this problem in a following
study.

%see also
%\S\ref{loc_iso_eos}
%We thus expect the standard Toomre
%parameter to be modified to $Q = c_s\sqrt{1-\tepsilon}\Omega/\pi
%G\Sigma$.

\subsubsection{Rossby wave instability}
The Rossby wave instability (RWI) is a non-axisymmetric, 2D shear
instability that operates in thin disks when it has radial structure
\citep{lovelace99,li00}. These studies consider adiabatic pure gas and 
show instability is possible if there is an extrema in the quantity
\begin{align}
 \eta = \frac{\kappa^2}{2\Omega\Sigma}\mathcal{S}^{2/\gamma},  
\end{align} 
where $\mathcal{S} = \Pi/\Sigma^\gamma$. Here, $\Sigma$ is the total
surface density and $\Pi$ is the vertically integrated pressure.

From the dusty/adiabatic gas equivalence, we deduce the corresponding
condition for RWI in a (locally isothermal or polytroipc) dusty gas is
also given by extrema in $\eta$ with $\mathcal{S} = \exp{s}$ (and
$\tepsilon\to \Sigma_\mathrm{d}/\Sigma$). Since
the dusty entropy $s$ reflects the dust distribution, we expect that
RWI may also be triggered by extrema in  the dust-to-gas ratio.

%gap edges. thin dust rings.
